---
title: "Coursera Developing Data Products"
subtitle: "Course Project"
author: "Timo Grossenbacher"
date: "October 25, 2015"
output: html_document
runtime: shiny
---

```{r, echo=FALSE, warning=FALSE, message=FALSE}
# load necessary packages
if(!require(dplyr)) {
  install.packages("dplyr", repos="http://cran.us.r-project.org")
  require(dplyr)
}
if(!require(tidyr)) {
  install.packages("tidyr", repos="http://cran.us.r-project.org")
  require(tidyr)
}
if(!require(ggplot2)) {
  install.packages("ggplot2", repos="http://cran.us.r-project.org")
  require(ggplot2)
}
if(!require(magrittr)) {
  install.packages("magrittr", repos="http://cran.us.r-project.org")
  require(magrittr)
}
if(!require(readxl)) {
  devtools::install_github("hadley/readxl")
  require(readxl)
}
if(!require(shinyapps)) {
  devtools::install_github("rstudio/shinyapps")
  require(shinyapps)
}
if(!require(BH)) {
  install.packages("BH", repos="http://cran.us.r-project.org")
  require(BH)
}
if(!require(lazyeval)) {
  install.packages("lazyeval", repos="http://cran.us.r-project.org")
  require(lazyeval)
}

if(!require(RColorBrewer)) {
  install.packages("RColorBrewer", repos="http://cran.us.r-project.org")
  require(RColorBrewer)
}

```

```{r, echo=FALSE, warning=FALSE, message=FALSE, results = 'hide'}
# this subsetting with bugfix in readxl
parties <- read_excel("input/parteienstaerke_mod_2.xlsx", sheet = 4)[1:24,] # same here
# all years 
years <- seq(1971, 2011, 4)
length(years) # we have 11 different time stamps
# read in input file (xlsx) - need to load a slightly modified input file as the original xlsx makes it really complicated
prepareYearlyData <- function(subsetDf, year){
  # add columns for numbers which are not contained in subsetDf
  subsetDf[,as.character(setdiff(parties$ID_BFS, as.numeric(names(subsetDf))))] <- NA
  # replace numerical column names with party names from parties
  partyNames <- paste(left_join(data.frame(ID_BFS = as.numeric(names(subsetDf))), parties, by = "ID_BFS")$Abbr_D, year, sep = "_")
  names(subsetDf) <- partyNames
  return(subsetDf)
}
prepareYearlyDataForVis <- function(subsetDf, year){
  # add columns for numbers which are not contained in subsetDf
  subsetDf[,as.character(setdiff(parties$ID_BFS, as.numeric(names(subsetDf))))] <- NA
  # replace numerical column names with party names from parties
  partyNames <- paste(as.numeric(names(subsetDf)), year, sep = "_")
  names(subsetDf) <- partyNames
  return(subsetDf)
}
raw_data_districts <- read_excel("input/parteienstaerke_mod_2.xlsx", sheet = 3, skip = 10, col_names = T)[1:147,] # replace this subsetting with bugfix in readxl
# doublecheck that "1" is contained as many times as we have years
table(names(raw_data_districts) == "1")[2] == length(years)
# this is true, so we can split the columns by this column name
indices <- which(names(raw_data_districts) == "1")
yearlyDataframes <- list()
for(i in 1:length(indices)){
  start <- indices[i]
  stop <- if(is.na(indices[i + 1])) ncol(raw_data_districts) else indices[i + 1] - 1 
  yearlyDataframes[[i]] <- prepareYearlyData(raw_data_districts[,start:stop], years[i])
}
# combine
filled_data_districts <- cbind(raw_data_districts[, 1:indices[1] - 1], yearlyDataframes)
# doublecheck: column numbers should be 2 + 11 * 24
dim(filled_data_districts)[2] == 2 + 11 * 24
# gather into long form
long_data_districts <- filled_data_districts %>%
  gather(ParteiJahr, Staerke,-c(BFSNr, BezName))
# doublecheck: number of rows should be 147 * 11 * 24
dim(long_data_districts)[1] == 147 * 11 * 24
# add new columns Jahr and Partei from splitted ParteiJahr
long_data_districts %<>% separate(col = ParteiJahr, into = c("Partei","Jahr")) %>%
  mutate(Jahr = as.Date(Jahr, "%Y")) %>%
  tbl_df()

# ontological choice: make NAs to 0es
long_data_districts[is.na(long_data_districts)] <- 0
```


# Swiss Party Strength Per District

The following application shows the party strengths for all major Swiss parties between 1971 and 2011 in all Swiss districts.

Federal elections happen all four years, thus for example in 1971 or 1975, etc. 

You have the possibility to chose 

 * a district -> only results for the selected district will be shown
 * a time range -> only results for the elections in the selected time range will be shown

```{r, echo=FALSE, warning=FALSE, message=FALSE}
distinct_colors <- c("#89C5DA", "#DA5724", "#74D944", "#CE50CA", "#3F4921", "#C0717C", "#CBD588", "#5F7FC7", 
"#673770", "#D3D93E", "#38333E", "#508578", "#D7C1B1", "#689030", "#AD6F3B", "#CD9BCD", 
"#D14285", "#6DDE88", "#652926", "#7FDCC0", "#C84248", "#8569D5", "#5E738F", "#D1A33D")

inputPanel(
  selectInput("district", label = "District", choices = unique(long_data_districts$BezName), selected = "Bezirk Solothurn"),
  sliderInput("yearrange", "Time range:",
                min = 1971, max = 2011, value = c(1971,2011), step = 4, sep = "")
)

renderPlot({
print(input$yearrange)
  startyear = as.Date(as.character(input$yearrange[1]), format = "%Y", origin = "1971-01-01")
  endyear = as.Date(as.character(input$yearrange[2]), format = "%Y", origin = "1971-01-01")

  bezirk <- long_data_districts %>%
    filter(BezName == input$district) %>% 
    filter(Jahr >= startyear & Jahr <= endyear)
  
  ggplot(bezirk, aes(x = Jahr, y = Staerke, color = Partei)) +
    geom_line(size = 2) +
    ylab("Party strength (%)") +
    xlab("Year") +
    theme_bw() +
    ggtitle(bezirk$BezName) +
    scale_color_manual(name = "Party", values = distinct_colors) +
    guides(color = guide_legend(ncol = 2))
})
```